Internet of Things: a Real-World Deployment Towards Smart Grids Integration

### 2013-07-02 doc/thesis/report.tex
implementation and appendices done

\documentclass[oneside,12pt,a4paper,final]{book}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{arabtex}
\usepackage{utf8}
\newcommand{\HRule}{\rule{\linewidth}{0.6mm}}
%\usepackage{hyperref}
\usepackage[acronym]{glossaries} 
\makeglossaries
\author{Nacer Khalil}
\title{Internet of Things: a Real-World Deployment Towards Smart Grids Integration}

@@

\section{Mote IPv6 backstation}
\paragraph{} 
This component that exist in one special mote in the \glsreset{wsn}\gls{wsn}. This mote is able to communicate with the rest of the \gls{wsn} using Zigbee and also with the external world using \gls{usb}. This make this mote a sink and a router because it has two interface cards. The software component deployed in this sink is a packet forwarder. This means that once it receives an \gls{ipv6} packet that is not in the \gls{wsn}, it is forwarded through \gls{usb}.
\paragraph{}
The mote's sole goal is forwarding packets in and outside of the \gls{wsn} while other sensor node do not need to worry about how the packet will arrive to its destination. This mote is part of the components aiming to solve the heterogeneity problem stated in the problem statement. By making the sensor nodes talk to any device as if it is part of the \gls{wsn}, they do not really have to worry about how it will arrive or even how did the packet get to them in the first place.

\section{IPv6 over USB tunnel}
\paragraph{}
The gateway server machine is equipped with two interface cards: the Wi-Fi network interface card and the Zigbee network interface card. The Zigbee network interface card is the one on focus in this section. To have a device communicating using Zigbee, one sensor node within the \gls{wsn} was used for this purpose and the software component implemented in this sensor node is the one described in section 4.6. That said, the packet is received through \gls{usb} but not read and recognized as a packet by the TCP/IP protocol stack. To do so, a tunnel had to be created for the purpose of extracting packets from the \gls{usb} port. Another goal is to forward \gls{ipv6} packets through that same \gls{usb} and this can only be done by the use of an \gls{ipv6} over \gls{usb} tunnel.

\section{Gateway packet transformation process}
The gateway packet transformation process is one of the main components that hides the heterogeneity from the rest of the system and the outside world. The

\chapter{Implementation}

\chapter{Findings}
\section{Experiment}
%\subsection{Methodology}
\section{Data gathering}
\section{Interpretation of result}

@@

\chapter{Future Work}

\appendix
\chapter{How to install the system}

@@

\singlespacing
\printglossaries
\bibliographystyle{IEEEtran} 
\bibliography{IEEEabrv,myrefs}

>>>>

\documentclass[oneside,12pt,a4paper,final]{book}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{arabtex}
\usepackage{utf8}
\newcommand{\HRule}{\rule{\linewidth}{0.6mm}}
%\usepackage{hyperref}
\usepackage{listings}
\lstset{language=C}
\usepackage[acronym]{glossaries} 
\makeglossaries
\author{Nacer Khalil}
\title{Internet of Things: a Real-World Deployment Towards Smart Grids Integration}

@@

\section{Mote IPv6 Backstation}
\paragraph{} 
This component that exist in one special mote in the \glsreset{wsn}\gls{wsn}. This mote is able to communicate with the rest of the \gls{wsn} using Zigbee and also with the external world using \gls{usb}. This make this mote a sink and a router because it has two interface cards. The software component deployed in this sink is a packet forwarder. This means that once it receives an \gls{ipv6} packet that is not in the \gls{wsn}, it is forwarded through \gls{usb}.
\paragraph{}
The mote's sole goal is forwarding packets in and outside of the \gls{wsn} while other sensor node do not need to worry about how the packet will arrive to its destination. This mote is part of the components aiming to solve the heterogeneity problem stated in the problem statement. By making the sensor nodes talk to any device as if it is part of the \gls{wsn}, they do not really have to worry about how it will arrive or even how did the packet get to them in the first place.

\section{IPv6 over USB Tunnel}
\paragraph{}
The gateway server machine is equipped with two interface cards: the Wi-Fi network interface card and the Zigbee network interface card. The Zigbee network interface card is the one on focus in this section. To have a device communicating using Zigbee, one sensor node within the \gls{wsn} was used for this purpose and the software component implemented in this sensor node is the one described in section 4.6. That said, the packet is received through \gls{usb} but not read and recognized as a packet by the TCP/IP protocol stack. To do so, a tunnel had to be created for the purpose of extracting packets from the \gls{usb} port. Another goal is to forward \gls{ipv6} packets through that same \gls{usb} and this can only be done by the use of an \gls{ipv6} over \gls{usb} tunnel.

\section{Gateway Packet Transformation Process}
The gateway packet transformation process is one of the main components that hide the heterogeneity from the rest of the system and the outside world. This component mainly deals with packet transformation which means that it extracts the data from the packet and create a completely new one that holds the data and forward it. This component is composed of three subcomponents:
\begin{itemize}
\item Packet sniffer
\item Packet transformer
\item Address translator
\end{itemize}
\paragraph{}
Figure \ref{fig:gateway_transformation} depicts the architecture of this component by showing the the different subcomponents that reside in this component and their interaction. It also shows the data flow within this component and what are the inputs and outputs of the system.

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.6]{img/packet_transformation.jpg}
\caption{Gateway packet transformation subsystem data flow}
\label{fig:gateway_transformation}
\end{figure}
\paragraph{}
The subsections 4.8.1 - 4.8.3 will go more into details in explaining the three subcomponents of the system.

\subsection{Packet Sniffer}
\paragraph{}
The packet sniffer subcomponent is the one responsible for sniffing and forwarding the packets to the component. It does not forward all the packets that are not going to the gateway machine. In other words, all the packets that are not meant to be received by the gateway server are sniffed and forwarded to the packet transformer that will continue the work.
\subsection{Packet Transformer}
\paragraph{}
Once the packet is received, the first thing this subcomponent does is check from where did it come, was it from the \gls{wsn} or the outside world. If it came from the outside world, the packet will only be examined if it is meant to go the \gls{wsn}. But an important and crucial problem presents itself:
\paragraph{Problem}
\underline{The outside world mostly uses \gls{ipv4} whereas the \gls{wsn} uses exclusively \gls{ipv6}} \\
Two networks using different network layer technology cannot communicate even though \gls{ipv6} is nothing more but a new version of the very known \gls{ipv4} and this is because \gls{ipv6} has no backward compatibility with \gls{ipv4}.
\paragraph{Solution}
To solve such a problem, we decide to provide all the nodes in the this system with both an \gls{ipv4} and \gls{ipv6} address. Some of the nodes have either \gls{ipv4} or \gls{ipv6} addresses, but for the technology missing we assign the missing one virtually. For example, a sensor node in the \gls{wsn} with a real \gls{ipv6} address equal to FEC0::1 will be assigned a virtual \gls{ipv4} address equal to 192.168.0.1 but he will not be aware of it. Only the Gateway packet transformation component will be aware of it and this is the mechanism used to transform the packets and hide the heterogeneity from the rest of the components of the system.
\paragraph{}
Now that a packet is received, it checks whether this component came from the \gls{wsn} or the outside world, depending on that, it extracts the data from the packet which can be a \gls{tcp} or \gls{udp} datagram. The source and destination addresses are extracted from the packet and translated to the other network technology. Afterwards, a new packet referring to the destination technology used is created by having as a source and destination addresses the translated addresses and as payload the the \gls{tcp} \gls{udp} datagram coming from the original packet. Finally, the packet is forwarded to the network.
\subsection{Address Translator}
The address translator is a very specialized subcomponent whose sole goal is to translate addresses between \gls{ipv4} and \gls{ipv6}. That is said, this component receives the address to translate and returns the translated one. Its concern is that the translation algorithm yields a bijective function which means that for any \gls{ipv4} address there is one and only one \gls{ipv6} address resulting from that source and vice versa. This statement compresses a set of statements which are the following:

\begin{itemize}
\item Any address belonging to one technology can be translated to the other technology. This is known as surjectivity.
\item Any address cannot be translated into more than one address from the other technology. This is known as injectivity.
\end{itemize}
\paragraph{}
The gateway packet transformation process is an important component responsible for dealing with the heterogeneity problem. It hides most of the heterogeneity in the system and make communication go independently of the data link layer or network layer technologies used.

\section{Sensor Data Storage Process}
\paragraph{}
The sensor data storage process is a component small component within the system whose sole purpose is to store such data. It seen by the \gls{wsn} as a TCP server, where sensor nodes get connected to it as soon as they are turned ON. The sensor node starts sending packet periodically (depending on the sampling rate specified) and whenever the sensor data is received, it is stored into a \gls{rdbms} where the schema is created as shown in figure \ref{fig:erd}.

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.5]{img/smart_grid_db.jpg}
\caption{Entity relationship diagram}
\label{fig:erd}
\end{figure}

\section{HTTP Sensor Data Sender} 
The HTTP sensor data sender is TCP server to which sensor nodes get connected to and start sending their data periodically. Once the data is received they are able to send data to the outside world periodically depending on the needs of the clients outside the system. This system counts on the deployment of an HTTP server at the level of outside the system client and send an HTTP request containing the sensor data received from the \gls{wsn}. 
\paragraph{}
The sensor nodes are able to send this data by themselves directly to the outside as this is part of the two-way communication but for convenience, it was decided that the \gls{wsn} sends its data to the gateway and it is up to the gateway to send such data or store it in the \gls{rdbms}.
\section{Mobile Client}
The mobile client is a component of the system but can be seen as an outside world client as it is here to receive sensor data and also send ON/OFF commands to the system.

\chapter{Deployment}
\paragraph{}
In this chapter, the test-bed will be discussed from the implementation point of view. That is said, each component within the system will be given as much details as possible in order for the reader to open the code and understand what is going. Chapter 4 discussed the components from the architectural aspect but no implementation and technology decisions were discussed and the goal of this chapter is to explain these details.
\section{Technology Enablers}
The system's implementation was built using a set of software and hardware technologies and standards that were used throughout the course of this project to make the test-bed achieve its objectives.
%\glsreset{wsn}
\subsection{Technologies used in \gls{wsn}}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.6]{img/crossbow.jpg}
\caption{Crossbow Professional Kit}
\label{fig:crossbow}
\end{figure}

The \gls{wsn} was built using Crossbow professional kit as shown in figure \ref{fig:crossbow} which counts 7 Micaz sensor nodes of type MPR2600 as shows in figure \ref{fig:micaz}.
\begin{figure}[htbp]
\centering
\includegraphics[scale=0.8]{img/micaz.png}
\caption{Crossbow MPR2600 Micaz Mote}
\label{fig:micaz}
\end{figure}

In addition, each mote is supplied with a sensor board of type Crossbow MTS400CA which contains temperature, humidity, barometric pressure, ambient light and infrared light sensors that are used to send different types of sensor data to the gateway. This board is shown in figure \ref{fig:mts400}.

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.5]{img/mts400ca.jpg}
\caption{Crossbow MTS400CA Sensor Board}
\label{fig:mts400}
\end{figure}
\paragraph{}
The WSN is run using \gls{tinyos} which is one of the most used opensource \gls{os} in \glspl{wsn}. Programs are implemented using \gls{nesc} which is the sole programming language used in the \gls{tinyos} environment. This programming language is event-oriented and its syntax is very influenced by C. In addition, C libraries can be used in \gls{nesc} programs.

\subsection{Technologies used in the Gateway}
\paragraph{}
The gateway's hardware machine is Dell Optiplex 9010 running a Linux environment more specifically the Ubuntu 13.04 distribution. The gateway has a set of technologies used to provide numerous services the system counts on. The programming languages used in the components hosted in this machine are C and \gls{j2se}. Netfilter is also used to control incoming and outgoing packets. Concerning the storage of sensor data MySQl is the \gls{rdbms} used.

\subsection{Technologies used in the Mobile Client}
\paragraph{}
The mobile used is an Android smart phone with Android 2.3. The application developed is using \gls{j2se}, PHP for database connectivity.

\section{Mote TCP Client And TCP Server}
Section 4.4 and 4.5 were dedicated to explain the TCP Client and TCP server separately. In this section, one section will explain both of them because from the implementation point of view,  both of them are implemented in the \gls{tinyos} program.
A typical \gls{tinyos} program has two parts: the components definition part and the module part. In the components definition part, the main components that will be used. These components refer to already implemented modules whose functionalities will be used in the in-development module. In the module part, we provide define the interfaces we will be using and the interfaces we will be providing if any \cite{ref14}. Back to definition part, after defining components we will be using, we then link the interfaces needed in the module to the components providing that interface. This is known as wiring.
The components needed in this program are the following:
\item MainC: provides the essential functionalities such as the initialization and closing of the program.
\item LedsC: provides access to the leds existing in the mote
\item Taos2550C: is a light sensor component that enable to read data
\item TimerMillic: A Timer that will be used for sampling
\item TCPSocket as client: Is the TCP client
\item TCPSocket as Server: Is the TCP server
In the modules part we define the components that will be used which are the following:
\item Boot: This is the interface that provides main functions of booting and closing
\item TCP as server: The main interface for TCP
\item TCP as client: The main interface for TCP
\item Leds: The interface to access the leds
\item timer as sensingTimer: The interface to program and use the timer
\item Read as Sensor: The interface to talk with the sensor
Some of the interface do not have all the functions provided and the not implemented function are implemented in the module. The following part will talk about each function that is to be implemented
\paragraph{After boot of the mote}
The first thing done after the mote is booted is to initialize the TCP client and server. The Client is given a random port by using the rand function. The TCP server is binded to port 7. The client is set to connect to the Data store in the gateway by providing the \gls{ipv6} address of the gateway and port number of the server. Afterwards, the timer is set to be triggered every 10 seconds which means that the sampling rate will be 360 sensor readings per hour. 
\paragraph{TCP Server data received}
Once the data is received from a client, it assumes that there are two data items separated by a space. The first item refers to the led number(1,2,3) and the second refers to the command 0 meaning turn Off and 1 means turn On. The string is tokenized and the data items are extracted and the command is executed accordingly.
\paragraph{Timer fired}
Once the timer is fired, this means it is time to read sensor data from the sensor. The command to read from the sensor is called which goes and reads data and once the data is read the event "read done" is triggered.
\paragraph{Read done}
Once the data is read and ready, the next thing is to transform it and then send it. Concerning the transformation, a series of equations are applied as seen in section 4.4 which build the data value from the original value gotten and it is ready to be sent. A message is created containing two values: the mote NODE\_ID which is the identifier of the mote predefined in the installation of the program in the mote and the sensor data value. These two values are separated by a space, encapsulated within a \gls{tcp} datagram and sent.
During the installation of the software the \glsreset{blip}\gls{blip} package should specified as it is a dependency of the software system. \gls{blip} provides the TCP/IP protocol stack that implements \gls{6lowpan} standard by \gls{ietf}. In  the installation the NODE\_ID is specified. If NODE\_ID is 1, this will make the mote have an \gls{ipv6} of FEC0::1, as the NODE\_ID is appended to the predefined prefix of \gls{wsn} during the installation.
\section{Mote IPv6 Backstation}
This component is provided as part of the \gls{blip} software package. It is \gls{tinyos} program whose goal is to forward packet back and forth between the \gls{wsn} and the outside. This mote is connected to the \gls{wsn} using Zigbee and to the outside world using \gls{usb}. This special mote has its \gls{ipv6} being FEC0::64. It receives \gls{ipv6} packets encapsulated in \gls{usb} frames, extracts them and encapsulates them in a Zigbee frame and vice versa.
During the installation of this \gls{tinyos} program into the mote sink, one should make sure that it is being installed in the mote sink as \gls{usb} is a necessity. The \gls{ipv6} address does not need to be provided as it is already predefined in the program.

\section{IPv6 over USB Tunnel}
\paragraph{}
This is a program that is installed in the gateway server machine. As the mote sink is connected to the gateway using \gls{usb}, the gateway will recognize it as a network interface card using \gls{ipv6} as a network technology. That is said, this component is part of the \gls{blip} software package and creates the network interface card module and starts extracting incoming \gls{ipv6} packets from the \gls{usb} frames and encapsulates outgoing \gls{ipv6} into \gls{usb} frames.
\paragraph{}
The installation and initialization of this component will be given more details in appendix A.
\section{Gateway Packet Transformation Process}
\paragraph{}
This component is in charge of hiding the heterogeneity of the system from the rest of the the components of the system and also the outside world. It does that by intercepting all the packet moving and transforms them according to the standards used by the recipient.
\paragraph{}
The program works with Netfilter in parallel. This means that Netfilter should drop all incoming and this program will take care of dealing with incoming packets separately. 
\paragraph{}
To make use of this process, there is an initial configuration that should be done to all the components of the system in order for the system to work. The gateway has two network interface cards: the Wi-Fi interface and the Zigbee interface provided through the tunnel. The first one is using \gls{ipv4} and the second is using \gls{ipv6}.
\paragraph{}
The program has three subcomponents as seen in section 4.8. The three components are the following:
\begin{itemize}
\item Packet sniffer
\item Packet transformer
\item Address translator
\end{itemize}
The following subsections will discuss each of these subcomponent separately from the implementation point of view.

\subsection{Packet sniffer}
\paragraph{}
The packet sniffer is the one in charge of sniffing packet from both network interface card, as Raw Frames. The code snippet shows the functions called to sniff the packets. We make use of the raw socket library that enables us to sniff data link layer frames before being dropped by Netfilter.
The main reason of dropping the packets by Netfilter is that the packets will not addressed to the computer and therefore the computer make send ICMP messages to the sender saying that the route is not the best and therefore the sender will never send packets again which is something we need to avoid to have.

\begin{lstlisting}
sockfd = socket( PF_PACKET, SOCK_PACKET , htons(ETH_P_ALL));
if (sockfd < 0)
	exit(1);

while(1)
{
	data_size = recvfrom(sockfd,buf,10000,0
		,(struct sockaddr *)&cliaddr,&clilen);
}

\end{lstlisting}
\paragraph{}
The first line of code is the initialization of the raw socket, we define the type, scope and name space of the packets to be sniffed. This will return a file descriptor that will allow us to call the socket. Afterwards, we start reading packets in an infinite loop.

\paragraph{}
After sniffing a packet, we must check if it is an \gls{ipv4} or \gls{ipv6} packet. The problem is that when reading from the Wi-Fi interface card, we get the whole frame where the network packet is encapsulated, but the Zigbee Tunnel only gives the raw \gls{ipv6} packet. Once a packet is read by the frame, from which interface card came is a problem and therefore a partial solution is found. We cast the packet to Wi-Fi frame and whenever we check for the network protocol, we get two solutions, 0 referring to \gls{ipv6} and 8 referring to \gls{ipv4}.
If we get zero, the whole data should be casted to the \gls{ipv6} data structure whereas if it is 8, we should move the pointer forward to drop the Wi-Fi header and cast the following part to the \gls{ipv4} data structure.
\paragraph{}
In each of those conditions we only select the TCP datagrams and this is why we check if the next header in the IP structure is equal to 6. After that we move the pointer to drop the IP header and only keep the TCP datagram that will have to be sent in a the newly transformed packet. Depending on the network technology of the sniffed packet, we should build a new packet with the technology that is recognized by the recipient.

\begin{lstlisting}
if(header->h_proto == 0) 
{
	if(ip6_hdr->ip6_nxt == 6)
	{
			buf += IP6_HDRLEN;
			... 			
			sendIPV4Packet(..)
	}
}
else if(header->h_proto == 8)
{
	if(ip_hdr->protocol == 6)
	{
		buf += (sizeof(struct ethhdr)+ IP4_HDRLEN);		
		...
		sendIPV6Packet(...)
	}
}
\end{lstlisting}

\subsection{Packet transformer}
\paragraph{}
The packet tranformer subcomponent is in charge of building two types of packets: \gls{ipv4} and \gls{ipv6} packets. As stated in section 4.8, all devices in this system have an \gls{ipv4} and \gls{ipv6} address where one of them is not recognized only at the gateway. As the \gls{wsn} excelusively supports \gls{ipv6} and they cannot be addressed without using their \gls{ipv6} address and any other device not using \gls{ipv6} may communicate with them by using their virtual \gls{ipv4} address at once the packets reach the gateway, it transform so that the data carried will be delivered and the packet will have instead of \gls{ipv4} source and destination addresses, \gls{ipv6} addresses that will be translated using the Address translator subcomponent.
\paragraph{}
This subcomponent is based on two important functions:
\begin{itemize}
\item sendIPv6Packet(..)
\item sendIPv4Packet(..)
\end{itemize}
\paragraph{}
The idea is the same for except that each technology has its own details such as filling different fields and computing the checksum. Another difference is when creating the \gls{ipv4} one should also create the Wi-Fi frame otherwise it won't be able to be sent through the corresponding network interface card whereas when building \gls{ipv6} packet one should just send the packet and the tunnel will take care of encapsulating it into a \gls{usb} frame.
The function receives the data to be sent which is in all cases a \gls{tcp} datagram along with the size of the datagram, the source and destination addresses that are already translated. Afterwards, the transformator creates the packet appends the datagram, fills the source and destination address and create a socket attached to the corresponding interface card and send the packet.
\begin{lstlisting}
int sendIPV6Packet(char* buffer,int size,char *src ,char *dest)
{
	//CREATE PACKET WITH ALL ITS DETAILS
	...
	memcpy (frame + IP6_HDRLEN, buffer, size);
	sd = socket (PF_PACKET, SOCK_RAW, htons (ETH_P_ALL))) < 0);
	sendto (sd, rame, length...);
}
\end{lstlisting}

\subsection{Address translator}
\paragraph{}
The address translator is the subcomponent that assigns \gls{ipv4} and \gls{ipv6} to all devices in this system. For instance, sensor nodes in the \gls{wsn} have \gls{ipv6} addresses but this subcomponent assigns to them \gls{ipv4} addresses.
\paragraph{}
To do so, the translation algorithm is purely bijective. That is said, any \gls{ipv4} is translated to one and only one \gls{ipv6} and vice versa. For a mobile client with \gls{ipv4} address of 10.50.1.50 to talk with a mote whose \gls{ipv6} address is FEC0::1, the mobile client should be aware of its translated to \gls{ipv4} address which is 192.168.0.1. Once the packet is received and knowing it is not addressed to it. The source and destination \gls{ipv4} addresses are translated to \gls{ipv6} making the source address equal to 0A32::0132 and the destination FEC0::1. The \gls{ipv6} packet is created and reaches the destination which is the sensor node. Once the response is going back the scenario is the same.
\paragraph{}
The translation algorithm has been tested extensively to make sure it is bijective and that any address can be translated to one and no more and than one address. When configuring the system, one should make sure that there is no two devices that have the same address real or virtual. Any \gls{ipv6} and \gls{ipv4} should be unique throughout the whole system. The \gls{wsn} should be granted an interval of \gls{ipv4} address so that no one will be able to have them besides the \gls{wsn} that uses them virtually.

\chapter{Findings}

\section{Experiment}
%\subsection{Methodology}
\section{Data Gathering}
\section{Interpretation of Results}

@@

\chapter{Future Work}
\paragraph{}
This thesis project's was to establish the infrastructure for integrating smart home within \glsreset{sg}\gls{sg} and make it part of the \glsreset{iot}\gls{iot}. This project enabled the deployment of the two-way communication as an infrastructure that can be used to build the \glsreset{dr}\gls{dr}. \gls{dr} is one of the most important topics in smart grids. It enables the utility and households to communicate with a set of services that any of them may request from the other. The current state of the art does not incorporate this communication as part of the smart home systems \cite{ref15}. But as there is a need for this communication, \gls{dr} will have to be implementation as part of the

@@

\appendix
\chapter{How to install the system}
\section{Hardware needed}
In order for the user to install the system, one should make sure he/she is having a computer that will be used for the sole purpose of hosting different services. The crossbow professional kit is a necessity and a router must be available to connect the different components through Wi-Fi. A smartphone or a separate computer with Android emulator may de used.

\section{Software installation}

\subsection{Operating system}
The machine should run Linux Ubuntu distribution (11.10 or more recent). You way use Windows as long as you Install a Virtual machine using Cygwin or other tools
\subsection{TinyOs Installation}
\begin{enumerate}
\item Execute: sudo gedit /etc/apt/source.list
\item Add to the opened file: deb http://hinrg.cs.jhu.edu/tinyos hardy main
\item sudo apt-get update 
\item sudo apt-get install tinyos-2.1.1
\item sudo gedit ~/.bashrc
\item add the following line: \\
export TOSDIR=\$TOSROOT/tos \\
export CLASSPATH=\$TOSROOT/support/sdk/java/tinyos.jar:.\$CLASSPATH \\
export MAKERULES=\$TOSROOT/support/make/Makerules \\
export PATH=/opt/msp430/bin:\$PATH \\
source /opt/tinyos-2.1.1/tinyos.sh \\

\item sudo chmod 777 \$TOSROOT
\end{enumerate}

\subsection{Configuring \gls{blip}}
\glsreset{blip}\gls{blip} should be compiled before being used. To do so, follow these instructions:
\begin{enumerate}
\item cd \$TOSROOT/support/sdk/c/sf \\
./bootstrap \\
./configure \\
make 
\item cd \$TOSROOT/support/sdk/c/blip \\
./bootstrap.sh \\
./configure \\
make 
\end{enumerate}

\subsection{Install Software in Motes}
The first thing to do is to install the backstation within the Mote sink (the mote with the \gls{usb} IO) To install it we do the following:
\begin{enumerate}
\item Plug the mote and type the command dmesg to see what \gls{usb} ID was assigned to it. It will generally be USB0.
\item cd \$TOSROOT/apps/IPBaseStation
\item make micaz blip install  mib520, /dev/ttyUSB0
\end{enumerate}
The backstation is now installed. We will now install the Mote TCP Client and Server in each sensor node. To do it, we should first download the code from the SVN repository.
\begin{enumerate}
\item Make sure you have SVN by typing svn --version
\item If you do not have SVN type sudo apt-get install subversion
\item download the file:
\item cd
\item svn checkout http://grid-aui-tinyos.googlecode.com/svn/trunk/ grid-aui-tinyos
\item cd grid-aui-tinyos//programming/TCPEcho
\end{enumerate}
Now take the MIB520 programming card, plug it and type dmesg to see where it plugged. lets call the location *USB. Now plug each mote in the MIB520 card and for each plugged mote do the following:

\begin{enumerate}
\item  make micaz blip install.*NUM mib520,/dev/tty*USB
\item make clean
\end{enumerate}
Note that *NUM refers to the NODE\_ID. For the first mote choose 1, for the second choose 2...

\subsection{Install Packet Transformation process}
To install this component, you must do the following:
\begin{enumerate}
\item Make sure you have CMake, type sudo apt-get install cmake
\item cd
\item grid-aui-tinyos//programming/netfiltering
\item cmake CMakeLists.txt
\item make
\end{enumerate}

\subsection{Install Gateway Sensor Data TCP Server}
To install the server you must do the following:
\begin{enumerate}
\item cd
\item grid-aui-tinyos//programming/GatewayServer/src
\item make
\end{enumerate}

\chapter{Run the System}
There is a set of instructions to follow in order as some of the components must be alive only when others are already alive. To run the system, do the following:
\begin{enumerate}
\item To turn on the Tunnel do the following
\subitem cd \$TOSROOT/support/sdk/c/blip/ 
\subitem sudo driver/ip-driver /dev/ttyUSB0 micaz. \\
 You must check that the mote sink is attached to USB0 by using dmesg, if not change it accordingly. After that, you should have an output that looks similar to the figure \ref{fig:tunnel_output}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.9]{img/tunnel_output.jpg}
\caption{IPv6 over USB tunnel execution output}
\label{fig:tunnel_output}
\end{figure}

\item To Run the Gateway Sensor Data TCP server do the following:
\subitem cd
\subitem cd grid-aui-tinyos/programming/GatewayServer/src
\subitem java Server

\item To Enable IP forwarding do the following:
\subitem sudo bash
\subitem echo 1 \textgreater /proc/sys/net/ipv4/ip\_forward
\subitem exit

\item  To drop all incoming packets: sudo iptables -P INPUT DROP
\item To disable ICPMP REDIRECT TO HOST:
\subitem sudo bash
\subitem /sbin/sysctl -w net.ipv4.conf.all.accept\_redirects=0
\subitem /sbin/sysctl -w net.ipv4.conf.all.send\_redirects=0
\subitem /sbin/sysctl -w net.ipv6.conf.all.accept\_redirects=0
\subitem /sbin/sysctl -w net.ipv6.conf.all.send\_redirects=0
\subitem exit

\item Run the Packet transformation process
\subitem cd
\subitem cd grid-aui-tinyos/programming/netfiltering
\subitem sudo my\_sniffer

\item Turn on the wireless sensor nodes one by one
\end{enumerate}

\paragraph{}
You should be able to get sensor data periodically. The mobile client access for turning On and Off will be explained in the Appendix C.

\singlespacing
\bibliographystyle{IEEEtran} 
\bibliography{IEEEabrv,myrefs}
